中国营商环境的地区差异、时空演变、收敛性与空间集聚性研究

### 2021-05-01 paper.Rmd
最近更新

load("final_index.rda")
final_index %<>% mutate(index = (index-min(index))/(max(index)-min(index)))

final_index %>% 
  dplyr::filter(year == 2019) %>% 
  group_sort(location,province,index) %>% 
  ggplot(aes(province,index,fill = location))+
  geom_col(color = alpha("white",0.5))+coord_flip()+
  # geom_hline(yintercept = qqq[[2]],linetype = "dashed",color = "black")+
  # geom_hline(yintercept = qqq[[3]],linetype = "dashed",color = "black")+
  # geom_hline(yintercept = qqq[[4]],linetype = "dashed",color = "black")+
  labs(x = "",y = "得分",fill = "地区")+
  scale_fill_manual(values = c(alpha("black",0.2),"darkgrey",alpha("black",0.6)))+
  # scale_color_viridis_d()+
  scale_y_continuous(limits = c(0,1.25))+
  theme(
    text = element_text(family = cnfont),
    plot.margin = margin(20,20,20,20),
    legend.position = "bottom",
    axis.text.y.left = element_text(size = 8)
  )

@@

```{r echo=FALSE}
load("mvb.rda")
mvb %>% pivot_wider(names_from = period,values_from = c(marginal,t)) %>% 
  set_names(c("省市","地区","第一期边际垂直beta","第一期半期收敛系数",
              "第二期边际垂直beta","第二期半期收敛系数")) %>% 
  select(-地区) %>% 
  arrange(第一期边际垂直beta) %>% 
  select(1,2,4,3,5) %>% 
  mutate(
    across(
      where(is.numeric),
     ~ round(.x,3)
    )
  ) %>% knitr::kable()
```

```{r echo=FALSE}
load("mvb.rda")
mvb %>% 
  group_sort(location,province,marginal) %>% 
  ggplot(aes(marginal,province,fill = location))+
  geom_col()+scale_colour_ggthemr_d()+
  # geom_text_repel(size = 3)+
  theme(
    text = element_text(family = cnfont),
    plot.margin = margin(20,20,20,20),
    legend.position = "bottom",
    axis.text.y.left = element_text(size = 8)
  )+
  scale_fill_manual(values = c(alpha("black",0.2),"darkgrey",alpha("black",0.6)))+
  labs(x = "",y = "",fill = "地区")+
  facet_wrap(.~period,scales = "free_x") 
```

>>>>

library(ggchicklet)
load("final_index.rda")
final_index %<>% mutate(index = (index-min(index))/(max(index)-min(index)))

final_index %>% 
  dplyr::filter(year == 2019) %>% 
  group_sort(location,province,index) %>% 
  ggplot(aes(province,index,fill = location))+
  geom_chicklet(radius = unit(2,"pt"),
                color = alpha("white",0.5))+
  coord_flip()+
  # geom_hline(yintercept = qqq[[2]],linetype = "dashed",color = "black")+
  # geom_hline(yintercept = qqq[[3]],linetype = "dashed",color = "black")+
  # geom_hline(yintercept = qqq[[4]],linetype = "dashed",color = "black")+
  labs(x = "",y = "得分",fill = "地区")+
  scale_fill_manual(values = c(alpha("black",0.2),"darkgrey",alpha("black",0.6)))+
  # scale_color_viridis_d()+
  scale_y_continuous(limits = c(0,1.25))+
  theme(
    text = element_text(family = cnfont),
    plot.margin = margin(20,20,20,20),
    legend.position = "bottom",
    axis.text.y.left = element_text(size = 8)
  )

@@

```{r echo=FALSE}
load("beta_reg.rda")
beta_reg %>% knitr::kable()
```

```{r echo=FALSE}
#####################################################################
# 动态展示图
load("mvb.rda")

group_sort <- function(df,group_var,id_var,target_var) {
  group_var <- enquo(group_var)
  id_var <- enquo(id_var)
  target_var <- enquo(target_var)

  map_df(
    df %>% count(!!group_var) %>% pull(1),
    ~ df %>% dplyr::filter(!!group_var == .x) %>% arrange(!!target_var)
  ) %>%
    mutate(!!id_var := fct_inorder(!!id_var,ordered = T))
} # end of "group_sort"

# 2010~2014年柱状图

mvb %>% dplyr::filter(period == "2010~2014年") %>% 
  group_sort(location,province,mar) %>% pull(1) -> lev

mvb %>% 
  mutate(
    province = factor(province,levels = lev)
  ) %>% dplyr::filter(period == "2010~2014年") %>% 
  ggplot(aes(province,mar,fill = location))+
  # geom_chicklet(radius = unit(2,"pt"),position = "dodge")+
  geom_col(position = "dodge")+
  scale_colour_ggthemr_d()+
  scale_x_discrete(labels = function(x) str_wrap(x,2))+
  theme(
    text = element_text(family = cnfont),
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5),
    axis.text.x.bottom = element_text(size = 8))+
  scale_fill_manual(values = c(
    alpha("black",0.2),"darkgrey",alpha("black",0.6)
  ))+
  labs(x = "",y = "",fill = "地区",title = "2010~2014年") -> p1

#####################################################################

# 2015~2019年柱状图

mvb %>% dplyr::filter(period == "2015~2019年") %>% 
  group_sort(location,province,mar) %>% pull(1) -> lev

mvb %>% 
  mutate(
    province = factor(province,levels = lev)
  ) %>% dplyr::filter(period == "2015~2019年") %>% 
  ggplot(aes(province,mar,fill = location))+
  # geom_chicklet(radius = unit(2,"pt"),position = "dodge")+
  geom_col(position = "dodge")+
  scale_colour_ggthemr_d()+
  scale_x_discrete(labels = function(x) str_wrap(x,2))+
  theme(
    text = element_text(family = cnfont),
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5),
    axis.text.x.bottom = element_text(size = 8))+
  scale_fill_manual(values = c(
    alpha("black",0.2),"darkgrey",alpha("black",0.6)
  ))+
  labs(x = "",y = "",fill = "地区",title = "2015~2019年") -> p2

library(patchwork)
(p1+p2)+plot_layout(guides = 'collect')&theme(legend.position = "bottom")

解释区域下降时使用“虹吸效应”。

```{r echo=FALSE}
load("mvb.rda")

group_sort <- function(df,group_var,id_var,target_var) {
  group_var <- enquo(group_var)
  id_var <- enquo(id_var)
  target_var <- enquo(target_var)

  map_df(
    df %>% count(!!group_var) %>% pull(1),
    ~ df %>% dplyr::filter(!!group_var == .x) %>% arrange(!!target_var)
  ) %>%
    mutate(!!id_var := fct_inorder(!!id_var,ordered = T))
} # end of "group_sort"

mvb %>% 
  dplyr::filter(period %in% c("2010~2014年","2015~2019年")) %>% 
  select(1:4) %>% pivot_wider(names_from = period,values_from = mar) %>% 
  mutate(gap = `2015~2019年` - `2010~2014年`) %>% 
  arrange(abs(gap)) %>% select(1:2,5) %>% pull(1) -> lev

mvb %>% select(1:4) %>% 
  mutate(province = factor(province,levels = lev)) %>% 
  pivot_wider(names_from = period,values_from = mar) %>% 
  mutate(tag = (`2015~2019年`>= `2010~2014年`)) %>% 
  dplyr::filter(province != "贵州") %>% 
  ggplot()+
  geom_point(aes(`2010~2014年`,province,shape = location),
             color = "lightgrey",size = 3)+
  geom_point(aes(`2015~2019年`,province,shape = location),
             color = "darkgrey" ,size = 3)+
  geom_segment(
    aes(x=`2010~2014年`,xend = `2015~2019年`,y = province,yend = province,
        linetype = tag),
    show.legend = F,
    arrow = arrow(length = unit(0.05, "inches"),type = "closed",angle = 25))+
  geom_vline(xintercept = 0,color = "black",linetype = "dotted")+
  scale_colour_ggthemr_d()+coord_flip()+
  theme(
    text = element_text(family = cnfont),
    plot.margin = margin(20,20,20,20),
    legend.position = "bottom"
  )+
  scale_y_discrete(labels = function(x) str_wrap(x,2))+
  scale_x_continuous(n.breaks = 10)+
  labs(x = "",y = "",shape = "地区")
```
